---
title: "Discussion 1"
author: "Student's Name Here"
output: html_document
---

```{r setup, include=FALSE}
# We will usually provide a chunk like this for you. You shouldn't change this chunk until and unless you know what you are doing.

# This is setting "global chunk options":

#   echo = TRUE causes every chunk to appear in the knitted file.
#   warning = FALSE causes warning messages to NOT appear in the knitted file.
#   message = FALSE causes other messages to NOT appear in the knitted file.

knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Preliminaries

- This file should be in `STAT240/discussion`.
- Set `STAT240/discussion` as your working directory. (This will be in problem 5).

The goal of today's discussion will be to review the syllabus, think about your learning goals for STAT 240, setting your directory and learn how to knit an Rmarkdown file to html.


# Problem 1

What are the dates for the first midterm, second midterm, and final exam for this course? As a reminder, makeup dates or times for the final exam are rarely offered and require extenuating or emergency circumstances.

> Write your answers here!


# Problem 2

What are your goals for taking this course and what are you hoping to learn?

> Write your answers here!

# Problem 3

Please ensure that the `STAT240` folder on your computer is set up according to the image attached in the assignment.

# Problem 4

In line 49 below, remove the text and add an R chunk and compute three times exp(10) plus five.

*You may find [Section 2.3.2 of the textbook ](https://bwu62.github.io/stat240-revamp/rstudio-intro.html#r-math) helpful.*

Add an R chunk here!

Ensure that the answer is appearing in the knitted file correctly. Do not just type the answer, write the code to get the answer.

# Problem 5 <b>(Working directory check)</b>

```{r}
# no written response needed, simply knit the document and
# check the output, this chunk should print:

getwd()
```

Copy the output of the previous line of code and change it to direct it to `Desktop/STAT240/discussion` and give it as input for `setwd()`. For example:

In my computer getwd() gives `"/Users/s.sahifa98/Downloads"`, then I change it to `"/Users/s.sahifa98/Desktop/STAT240/discussion"` and put it inside `setwd("/Users/s.sahifa98/Desktop/STAT240/discussion")`.

Run getwd() again to make sure it's set correctly.

# Submission

When you have answered the above questions, please knit this file and submit **both** the R Markdown file (.Rmd) and the knitted html file (.html).

We will ask for both files for every assignment in this course.

We will grade directly from the .html file, so make sure you are satisfied with how it looks before you submit. We ask for your .Rmd as a backup in case your .html file has errors in it. 
