---
title: "Basic R stuff - 002"
author: "Sahifa"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE)
```

# Wednesday 09/09
# Basic R stuff

## Mathematical Operations:

+, -, /, *, ** or ^ (raised to the power of), %% (remainder)

```{r}
3/2
3*2
3**2
3^2
3%%2
```

# Logical Operations

<, >, <=, >=, == (equivalency), !=

```{r}
3 >= 3
3 < 2
1 == -1
0 == NaN # coming up soon
```

`and` operator `&`, `or` operator `|`

```{r}
T & T
T | F
T & F
```



## Variables

format: name = value, name <- value

```{r}
a <- 1
a
```

Rules:

1) All names should start with letters, you can use numbers and symbols after a letter in names of variables

Eg: 1a

```{r}
1a = 2
a1 = 2

a1 # calling a variable
```

2) Names are case sensitive

```{r}
A = 1
a1_ = 2
a1* = 2
a1. = 2
```

3) Any weird names should be enclosed in quotations. Calling weird named variables require you to enclose the variable name in backticks ``.

```{r}
'shop name' = "Starbucks"
'1a' = 1

'1a' = 10

`1a`
'1a'
```


### Removing defined variables from environment

```{r}
remove(`1a`)
```


# Vectors

all numbers between 1 to 5 as numbers.

- `c()`: stands for combine

```{r}
numbers = c(1, 2, 3, 4, 5)
```

```{r}
3+2
```

# Friday 09/11

### Types of variables

- Numbers: These are called numeric variables. dbl, int, num, date, etc.

Check the type of variable using the function `class()`

```{r}
a

class(a)
```

- characters: these are texts, are defined inside quotations.

```{r}
name = "Sahifa"
my_name = 'Sahifa'

name == my_name

no. = "22"
no.+1

num = as.numeric(no.)
class(num)

as.numeric(name)
```

- Logical values: TRUE, FALSE, NA (stands for `not available` and is interpreted as missing value), NaN (`Not A Number`)

```{r}
TRUE
FALSE
T
F

T+2
F/1

0/0
NA+1

NA == NA
```


# Vectors

all numbers between 1 to 5 as numbers.

- `c()`: stands for combine

```{r}
numbers = c(1, 2, 3, 4, 5)
numbers
```

```{r}
nu_chr = c(1,2, 3, "four", "five", "name")
num_log = c(1, 2, T, FALSE, 40)
log_chr = c(T, TRUE, FALSE, "F", "100")

num_log
log_chr
```


- `:` this is used to create sequences of numbers with a diff of one

format: a:b, where a and b are numbers

```{r}
1:20
pi:20
-10:10

10:1
```

- `seq()`:

```{r}
?seq

seq(from = 10, to = 1, by = -0.1)

seq(1, 10, length.out = 98)
```

- `rep()`: repetition

```{r}
?rep

rep(numbers, times = 10)
```


### Operations with vectors:

Mathematical opeartions:

```{r}
a
b = c(1, 11, 21)
b
numbers
```

```{r}
a + b

b+numbers
```

1+1
11+2
21+3
1+4
11 + 5






