---
title: "Basic R stuff - 002"
author: "Sahifa"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE)
```


# Basic R stuff

## Mathematical Operations:

+, -, /, *, ** or ^ (raised to the power of), %% (remainder)

```{r}
3/2
3*2
3**2
3^2
3%%2
```

# Logical Operations

<, >, <=, >=, == (equivalency), !=

```{r}
3 >= 3
3 < 2
1 == -1
0 == NaN # coming up soon
```

`and` operator `&`, `or` operator `|`

```{r}
T & T
T | F
T & F
```



## Variables

format: name = value, name <- value

```{r}
a <- 1
a
```

Rules:

1) All names should start with letters, you can use numbers and symbols after a letter in names of variables

Eg: 1a

```{r}
1a = 2
a1 = 2

a1 # calling a variable
```

2) Names are case sensitive

```{r}
A = 1
a1_ = 2
a1* = 2
a1. = 2
```

3) Any weird names should be enclosed in quotations. Calling weird named variables require you to enclose the variable name in backticks ``.

```{r}
'shop name' = "Starbucks"
'1a' = 1

'1a' = 10

`1a`
'1a'
```


### Removing defined variables from environment

```{r}
remove(`1a`)
```


# Vectors

all numbers between 1 to 5 as numbers.

- `c()`: stands for combine

```{r}
numbers = c(1, 2, 3, 4, 5)
```

```{r}
3+2
```





