---
title: "dplyr-002"
author: "Sahifa"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = T)
library(tidyverse)
```

#dplyr

- Pipe operator:`%>%`

`|>`

Windows: ctrl+shift+M
Mac: cmd+shift+M

$$log(sqrt(100^3))$$

```{r}
log(sqrt(100^3))

logb(sqrt(100^3), base = 10)

100^3 %>% sqrt() %>% logb(base = 10)
```

h(g(f(x, y), z, p), a, c)

x %>% f(y) %>% g(z, p) %>% h(a,c)

p %>% g(z) = g(p,  z) not equal to g(z, p)

# dplyr functions

```{r}
grocery_data = read_csv("../data/grocery_data.csv")
grocery_list = read_csv("../data/grocery_list.csv")
```

### column-wise functions:

- `mutate()`: add a column to your dataset. arguments will follow tibble format, `column name` = values that go in given as a vector

```{r}
updated_grocery = grocery_data %>% mutate(tax = price*0.02, final_price = price+tax)
```

- `select()`: this functions will create a dataset with only the lsited columns

```{r}
updated_grocery %>% select(item, type, final_price)

updated_grocery %>% select(item:tax)
```

- `relocate()`: moves around columns in your dataset

```{r}
updated_grocery %>% relocate(item, type, price, tax)


updated_grocery %>% relocate(type, .after = item)
updated_grocery %>% relocate(type, .before = price)
```

- `rename()`: renames existing columns in the dataset. Argument: `new name` = `old name`

```{r}
updated_grocery %>% rename("final price" = final_price) %>% select(item, `final price`)

updated_grocery %>% rename("final price" = final_price, ITEM = item)
```

### Row-wise functions

- `arrange()`: same as the sort function


```{r}
updated_grocery %>% arrange(price) %>% arrange(item)

updated_grocery %>% arrange(-price)

updated_grocery %>% arrange(desc(item))
```

# Friday 09/25

- `filter()`: filters out rows that satisfy a given condition. You provide this function with logical statements and then the output only contains the rows that were TRUE for that logical statement

```{r}
updated_grocery %>% filter(type == "vegetable") # gives rows where type is a vegetable

updated_grocery %>% filter(type == "fruit")

# get columns for all the items whose price is more than $3.

updated_grocery %>% filter(price > 3)

# filter function is the best to filter out NA values

updated_grocery[3, 2] = NA

updated_grocery

updated_grocery %>% filter(price == NA) # this one will not extarct NA because that logical statement returning NAs and not T or F


updated_grocery %>% filter(is.na(price))

x = NA

x == NA

updated_grocery$price == NA
is.na(updated_grocery$price)

updated_grocery %>% filter(!is.na(price))

# get columns for all the items whose price is not $3.

updated_grocery %>% filter(price != 3)


# assuming type which has NA values, and we already have price column with an NA, to remove both of those NAs we have to combine two differnt logical statemnets with `&` operator

updated_grocery %>% filter(!is.na(price) & !is.na(type))
```

- `slice_max()` and `slice_min()`

```{r}
updated_grocery %>% slice_min(price, n= 2)
```

### Group-wise functions

- `summarize()`: creates a summary table. This function needs to you use to other summary functions, for example: sd(), mean(), sum(), etc. Format of the argumnet: `name of col` = description of the col

```{r}
updated_grocery %>% summarise(max_price = max(price, na.rm = T), avg_price = mean(price, na.rm = T), min_price = min(price, na.rm = T), log_price = log(price)) # will not work with the later R versions because they have updated the tidyverse summarize to only work with summary functions.
```


- `group_by()`:

```{r}
updated_grocery %>% group_by(type) %>% slice_max(price, n = 2)

updated_grocery %>% group_by(type) %>% summarise(avg_price = mean(price, na.rm = T), sd_price = sd(price, na.rm = T))
```









