---
title: "dplyr-002"
author: "Sahifa"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = T)
library(tidyverse)
```

#dplyr

- Pipe operator:`%>%`

`|>`

Windows: ctrl+shift+M
Mac: cmd+shift+M

$$log(sqrt(100^3))$$

```{r}
log(sqrt(100^3))

logb(sqrt(100^3), base = 10)

100^3 %>% sqrt() %>% logb(base = 10)
```

h(g(f(x, y), z, p), a, c)

x %>% f(y) %>% g(z, p) %>% h(a,c)

p %>% g(z) = g(p,  z) not equal to g(z, p)

# dplyr functions

```{r}
grocery_data = read_csv("../data/grocery_data.csv")
grocery_list = read_csv("../data/grocery_list.csv")
```

### column-wise functions:

- `mutate()`: add a column to your dataset. arguments will follow tibble format, `column name` = values that go in given as a vector

```{r}
updated_grocery = grocery_data %>% mutate(tax = price*0.02, final_price = price+tax)
```

- `select()`: this functions will create a dataset with only the lsited columns

```{r}
updated_grocery %>% select(item, type, final_price)

updated_grocery %>% select(item:tax)
```

- `relocate()`: moves around columns in your dataset

```{r}
updated_grocery %>% relocate(item, type, price, tax)


updated_grocery %>% relocate(type, .after = item)
updated_grocery %>% relocate(type, .before = price)
```

- `rename()`: renames existing columns in the dataset. Argument: `new name` = `old name`

```{r}
updated_grocery %>% rename("final price" = final_price) %>% select(item, `final price`)

updated_grocery %>% rename("final price" = final_price, ITEM = item)
```

### Row-wise functions

- `arrange()`: same as the sort function


```{r}
updated_grocery %>% arrange(price) %>% arrange(item)

updated_grocery %>% arrange(-price)

updated_grocery %>% arrange(desc(item))
```










