---
title: "ggplot_002"
author: "Sahifa"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = T)
```
# Wednesday 09/16

```{r}
# Assuming your working directory is set to STAT240/Lecture Notes

library("tidyverse")
mendota = read_csv("../data/lake-mendota-winters-2025.csv")
glimpse(mendota)
```

# GGPlot - Grammar of Graphics

## Step- by - step process

```{r}
# creating a ggplot using mendota dataset and columns
ggplot(data = mendota) # step 1

ggplot(data = mendota, mapping = aes(x = year1, y = duration)) # step 2
```


- step 3 of ggplot: layers go on the ggplot

geom_functions

Examples:

- geom_point(): point graph
- geom_line(): line graph
- geom_bar(): bar graphs
- geom_histogram(): creates histograms
- geom_density(): creates density plots



```{r}
# step 3

ggplot(data = mendota, mapping = aes(x = year1, y = duration)) +
  geom_point(col = "green") + 
  geom_line(color = "seagreen") + 
  geom_smooth(col = "red")
```


# Aesthetics:

- Aes depending on the location of the input

### Global aesthetic: 

Any plot informtaion that goes inside of `ggplot()` is called a global aes and gets applied to all following `geom_` functions.


### Local aesthetic:

Any plot information inside of individual `geom_` functions is called a local aes and gets applied to only that `geom`


- Aes depending on the source of information

### Variable aesthetic: 

Anytime the informatin displayed on the graph is varying from point to point we cal it a variable aes

### Constant aesthetic

Information that doesn't change from point to point is what we call a constant aesthetic


```{r}
ggplot(data = mendota, mapping = aes(x = year1, y = duration)) +
  geom_point(mapping = aes(col = intervals)) + 
  geom_line(col = "seagreen")

ggplot(data = mendota, mapping = aes(x = year1, y = duration)) +
  geom_point(mapping = aes(col = as.factor(intervals))) + 
  geom_line(col = "seagreen")

x_axis = mendota$year1
y_axis = mendota$duration
col = mendota$intervals
ggplot(mapping = aes(x = x_axis, y = y_axis)) +
  geom_point(mapping = aes(col = col)) + 
  geom_line(col = "seagreen")
```

# Two variable plots

- `geom_point()` and `geom_line()`

- `geom_smooth()`:

# Friday 09/18

- `geom_smooth()`: trend line for your dataset

In this class, we are interested in linear regression

```{r}
# make sure you run the `library(tidyverse)` and mendota dataset line of code before starting

ggplot(data = mendota, mapping = aes(x = year1, y = duration)) +
  geom_point() + 
  geom_smooth(method = "lm", se = F)
```

- `geom_col()`: column graph

```{r}
decade = count(mendota, decade)
decade

ggplot(decade, aes(decade, n)) +
  geom_col(fill = "yellow", col = "pink")

# another example of a column graph which cannot be recreated with geom_bar()

summary_decade = mendota %>% group_by(decade) %>% summarise(avg = mean(duration)) # calculating the avg duration the lake was closed in each decade

summary_decade

ggplot(summary_decade, aes(decade, avg)) +
  geom_col()
```


# One variable plots

- `geom_bar()`: bar graph for categorical variables


```{r}
ggplot(mendota, aes(decade)) + 
  geom_bar(fill = "yellow", col = "pink")
```

- `geom_histograms()`:


```{r}
ggplot(mendota, aes(duration)) +
  geom_histogram(fill = "skyblue", col = "purple")
```

- `geom_density()`: density curve of the distribution of the variable

```{r}
ggplot(mendota, aes(duration)) +
  geom_density()
```

```{r}
ggplot(mendota, aes(duration)) +
  geom_density(col = "red", fill = "yellow") +
  geom_histogram(mapping = aes(y = after_stat(density)), alpha = 0.5)
```

- `geom_boxplot()`: this is a one- dimensional graph that displays the distribution of the variable

```{r}
ggplot(mendota, aes(duration)) +
  geom_boxplot(fill = "yellow")
```










