---
title: "Join and pivots (last topic for midterms)"
author: "Sahifa"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE)
library(tidyverse)
```
# Wednesday 09/30

A = grocery_list
B = grocery_data

```{r}
A = grocery_list
B = grocery_data

A; B
```

### Mutating joins

name_join(A, B, by = join_by())

- left_join(): give all rows from the first daatset or dataset A and onlyu the matching (with A) rows from B.
- right_join()
- inner_join()
- full_join()


A: 15 rows and 2 columns (1 column was `item`)
B: 12 rows and 3 columns (1 column was `item`)

Common rows in A and B is 9, extra rows in A is 6 (15-9), extra rows in B is 3 (12-9)

Output dataset: Columns: common column + leftover columns of dataset A + leftover columns of dataset B = 1 + 1 + 2 = 4 columns in total

Rows: 15 rows for all the entries in the common column in dataset A, 

```{r}
left_join(A, B) # rows = 15, columns = 4
```

```{r}
dim(A); dim(B)

right_join(A, B) # Rows = 12, columns = 4

left_join(B, A)
```

```{r}
# gives rows that are common in both the A and B 
inner_join(A, B) # rows = 9 (those are common items in both the A and B datsaets), columns = 4
```

```{r}
# gives all the rows from both the datasets

full_join(A, B) # rows = 18 (9+6+3), columns = 4

full_join(B, A)
```

### Filtering Joins: These joins only give you columns from the first dataset (dataset A).

- `semi_join()`: Gives you all rows from the first dataset that match the commom column entries in the second dataset and only give columns from the first dataset

- `anti_join()`: Gives you all rows from the first dataset that match the uncommom column entries in the second dataset and only give columns from the first dataset


```{r}
semi_join(A, B) # rows = 9, columns = 2
```

```{r}
anti_join(A, B)
```

```{r}
drinks = tibble(
  Row = 1:4,
  Drink = c("Milk Tea", "Espresso", "Latte", "Espresso"),
  Cafe = c("Happy Tea", "Daily Grind", "Daily Grind", "Bev's Beverages"),
  Price = c(4, 2, 3, 2)
)

drinks
```


```{r}
discounts = tibble(
  Row = 1:3,
  Coupon = c("25%", "50%", "5%"),
  Location = c("Happy Tea", "Daily Grind", "The Roast")
)

discounts
```


Dataset A: Columns = 4, Rows = 4
Dataset B: Columns = 3, Rows = 3

Common columns: 2 (Row, Cafe == Location)
Common entries/rows: 2 daily grind and happy tea (but dataset A has daily grind twice)
Uncommon rows: 1 in both

```{r}
left_join(drinks, discounts)

left_join(drinks, discounts, by = join_by(Cafe == Location))

left_join(drinks, discounts, join_by(Cafe == Location, Row))
```

```{r}
band_instruments
band_members
```









