---
title: "Basic R stuff - 003"
author: "Sahifa"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Basic R 

## Operators:

### Mathematical Operators:

+, -, *, /, ** or ^ (raised to power of), %% (remainder of a division)


```{r}
3+2
3/2
3%%2
3**2
3^2
```

### Logical Operation:

<, <=, >, >=, == (equivalency), !=

```{r}
3>2
3<=2
3==3
3!=3
```

'and' operator: `&`
'or' operator: `|`

```{r}
T|T
T|F
F|T
F|F
```

### Other mathematical calculation

```{r}
sqrt(2)
log(10)
exp(10)
```


# Variables:

format: name = value, name <- value


```{r}
a = 1 # defined a variable called `a` which has a value 1

a # calling a variable to get it's value
```

Rules:

1) Every variable name should start with letters, and not numbers or symbols. And it can be followed by numbers or symbols like `_` or `.`.

Eg: 1a shouldn't be a name for a variable

```{r}
1a = 10
*a 
_a = 10

a1 = 2
a_ = 2
```

2) Variable names are case sensitive

```{r}
A = 1
```

3) Any weird variable names should be enclosed in quotation.

```{r}
'my fav number' = 2
'1a' = 10

`1a`
`my fav number`
```


```{r}
a
```

### Types of variables

- numeric: 

Different ways R saves numbers: double, numeric, float, integer. BUt all of them are still considered numeric for our purposes

```{r}
class(a)
```

- character values:

texts, to define these variables you have to enclose the value in quotations.

```{r}
b = '22'
class(b)

b+1

'my name' = "Sahifa"

'my name with a space' = "Sahifa "

`my name` == `my name with a space`
```



