---
title: "dplyr-003"
author: "Sahifa"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# dplyr

- Pipe operator: ` %>% `

or `|>`

Keyboard shortcuts:
MAC: cmd+shift+M
Windows: ctrl+shift+M

log(sqrt(sum((c(1, 2, 4, 7))^2)))

```{r}
log(sqrt(sum((c(1, 2, 4, 7))^2)))

(c(1, 2, 4, 7))^2 %>% sum() %>% sqrt() %>% log()
```

logb(sqrt(sum((c(1, 2, 4, 7))^2)), base = 10)

```{r}
logb(sqrt(sum((c(1, 2, 4, 7))^2)), base = 10)

(c(1, 2, 4, 7))^2 %>% sum() %>% sqrt(2) %>% logb(base = 10)
```


x %>% f(y) %>% g(a, b, c) %>% h(p)

f(x, y) %>% g(a, b, c) %>% h(p)

g(f(x, y), a, b, c) %>% h(p)

h(g(f(x, y), a, b, c), p)

# dplyr functions

```{r}
# assuming that you have set your STAT240/lecture notes as your directory and those datasets are in STAT240/data

grocery_data = read_csv("../data/grocery_data.csv")
grocery_list = read_csv("../data/grocery_list.csv")
```


### column-wise functions

- `mutate()`: adds columns to your existing dataset

```{r}
(updated_grocery = grocery_data %>% mutate(tax = price*0.02, final_price = price+tax))
```

- `select()`: will choose/keep only the columns mentioned


```{r}
updated_grocery %>% select(item, final_price, type)

updated_grocery %>% select(item:tax)


# c(c1, c3, c7:c15, c20, c50)
```

- `relocate()`: 

```{r}
updated_grocery %>% relocate(item, type, price, tax)
updated_grocery %>% relocate(type, .after = item) 
updated_grocery %>% relocate(c(tax, final_price), .before = type)

?relocate
```

- `rename()`: rename pre-existing columns, format of arg: `new name` = `old name`

```{r}
updated_grocery %>% rename("final price" = final_price, Item = item)
```

### Row-wise functions

- `arrange()`: sort() equivalent for columns in a dataset

```{r}
updated_grocery %>% arrange(price)

updated_grocery %>% arrange(item)

updated_grocery %>% arrange(-price)

updated_grocery %>% arrange(desc(item))
```
# Friday 09/25

- `filter()`: filter out the rows depending on the logical statements.

```{r}
# filter out only the rows corresponding to "vegetable"
updated_grocery %>% filter(type == "vegetable")
updated_grocery %>% filter(type != "fruit")

# filter out items that priced more than $3.

updated_grocery %>% filter(price > 3)

updated_grocery[3,2] = NA

updated_grocery

# filter out the row that has missing value using dplyr

updated_grocery %>% filter(price == NA)

updated_grocery %>% filter(price > 0)

updated_grocery$price == NA

is.na(updated_grocery$price)

updated_grocery %>% filter(is.na(price))

# filter out the row that do not have missing value using dplyr

updated_grocery %>% filter(is.na(price) == FALSE)
updated_grocery %>% filter(!is.na(price))

updated_grocery[4, 5] = NA

updated_grocery

updated_grocery %>% filter(!is.na(price) & !is.na(final_price)) %>%  filter(price > 3)
```

- `slice_min()` and `slice_max()`:

```{r}
updated_grocery %>% slice_min(price) # least priced item

updated_grocery %>% slice_min(price, n = 3)
```

### Group-wise functions

- `summarize()`: create summary tables/datasets. Uses summary functions: sum(), mean(), sd(), median(), etc. Format of the arguments: `column name` = column description

```{r}
updated_grocery %>% summarise(min_price = min(price, na.rm = T), max_price = max(price, na.rm = T), avg_price = mean(price, na.rm = T))

updated_grocery %>% summarise(min_price = min(price, na.rm = T), max_price = max(price, na.rm = T), avg_price = mean(price, na.rm = T), log = log(price)) # will create error
```

- `group_by()`: 

```{r}
updated_grocery %>% group_by(type)

updated_grocery %>% select(-type) %>% filter(!is.na(price)) %>% group_by(price > 2) %>% summarise(total_price = sum(final_price, na.rm = T))


updated_grocery %>% group_by(type) %>% summarise(total_price = sum(final_price, na.rm = T))

# try group_by with all the function we have learned up until now

```

