---
title: "dplyr-003"
author: "Sahifa"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# dplyr

- Pipe operator: ` %>% `

or `|>`

Keyboard shortcuts:
MAC: cmd+shift+M
Windows: ctrl+shift+M

log(sqrt(sum((c(1, 2, 4, 7))^2)))

```{r}
log(sqrt(sum((c(1, 2, 4, 7))^2)))

(c(1, 2, 4, 7))^2 %>% sum() %>% sqrt() %>% log()
```

logb(sqrt(sum((c(1, 2, 4, 7))^2)), base = 10)

```{r}
logb(sqrt(sum((c(1, 2, 4, 7))^2)), base = 10)

(c(1, 2, 4, 7))^2 %>% sum() %>% sqrt(2) %>% logb(base = 10)
```


x %>% f(y) %>% g(a, b, c) %>% h(p)

f(x, y) %>% g(a, b, c) %>% h(p)

g(f(x, y), a, b, c) %>% h(p)

h(g(f(x, y), a, b, c), p)

# dplyr functions

```{r}
# assuming that you have set your STAT240/lecture notes as your directory and those datasets are in STAT240/data

grocery_data = read_csv("../data/grocery_data.csv")
grocery_list = read_csv("../data/grocery_list.csv")
```


### column-wise functions

- `mutate()`: adds columns to your existing dataset

```{r}
(updated_grocery = grocery_data %>% mutate(tax = price*0.02, final_price = price+tax))
```

- `select()`: will choose/keep only the columns mentioned


```{r}
updated_grocery %>% select(item, final_price, type)

updated_grocery %>% select(item:tax)


# c(c1, c3, c7:c15, c20, c50)
```

- `relocate()`: 

```{r}
updated_grocery %>% relocate(item, type, price, tax)
updated_grocery %>% relocate(type, .after = item) 
updated_grocery %>% relocate(c(tax, final_price), .before = type)

?relocate
```

- `rename()`: rename pre-existing columns, format of arg: `new name` = `old name`

```{r}
updated_grocery %>% rename("final price" = final_price, Item = item)
```

### Row-wise functions

- `arrange()`: sort() equivalent for columns in a dataset

```{r}
updated_grocery %>% arrange(price)

updated_grocery %>% arrange(item)

updated_grocery %>% arrange(-price)

updated_grocery %>% arrange(desc(item))
```




