---
title: "ggplot_003"
author: "Sahifa"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE)
```

```{r}
# Assuming that working directory is set to STAT240/Lecture Notes 

library(tidyverse)
mendota = read_csv("../data/lake-mendota-winters-2025.csv")
glimpse(mendota)
```

# GGplot2

GG: Grammar of Graphics

### Step - by - step process

```{r}
# Step 1

ggplot()

# Step 2

ggplot(data = mendota, mapping = aes(x = year1, y = duration))
```

- step 3: `geom_` functions

`geom_point()`: point graph
`geom_line()`: line graph
`geom_histogram()`: create histograms
`geom_density()`: creates a density curve
`geom_bar()`: bar graph


```{r}
# Step 3: layers

ggplot(data = mendota, mapping = aes(x = year1, y = duration)) + 
  geom_point(col = "red") + 
  geom_line(col = "blue")
```


```{r}
ggplot() + 
  geom_point(data = mendota, mapping = aes(x = year1, y = duration), col = "red") + 
  geom_line(col = "blue")
```


# Aesthetic

- classifictation depending on location of the information provided
### Global aesthetic: 
any info provided to `ggplot()` is called a global aesthetic and will affect all `geom_`functions following it.


### Local aesthetic:

any info provided to individual `geom_` functions will only be displayed in that geom_ output.

- depending on type of information

```{r}
ggplot(data = mendota, mapping = aes(x = year1, y = duration)) + 
  geom_point(mapping = aes(col = intervals)) + 
  geom_line(col = "blue")

ggplot(data = mendota, mapping = aes(x = year1, y = duration)) + 
  geom_point(mapping = aes(col = as.factor(intervals))) + 
  geom_line(col = "blue")
```

### Variable aesthetic: 

any information that changes from one point to the other is called a variable aesthetic

### Constant aesthetic:

any information that remains constant throughout the graph is called a constant aesthetic


# Two Variable plots

- `geom_point()` and `geom_line()`

- `geom_smooth()`:

```{r}
ggplot(data = mendota, mapping = aes(x = year1, y = duration)) + 
  geom_point(mapping = aes(col = as.factor(intervals))) +
  geom_smooth(method = "lm", se = F)

ggplot(data = mendota, mapping = aes(x = year1, y = duration)) + 
  geom_point(mapping = aes(col = as.factor(intervals))) +
  geom_smooth(method = "lm", se = F, mapping = aes(col = as.factor(intervals)))
``` 

# Friday 09/18

- `geom_col()`: Column/bar graphs uses both the x and y variable where the second variable is usually a summary of the data (first variable).

```{r}
decade_count = count(mendota, decade)
decade_count

ggplot(decade_count, aes(decade, n)) +
  geom_col(fill = "yellow", col = "red")

# an example that you can not recreate with geom_bar()

decade_summary = mendota %>% group_by(decade) %>% summarise(avg_duration = mean(duration))

decade_summary

ggplot(decade_summary, aes(decade, avg_duration)) +
  geom_col()
```

# One variable plots

- `geom_bar()`: bar graphs for a categorical varioable that displays the *count* of the variable


```{r}
ggplot(mendota, aes(decade)) +
  geom_bar()
```

- `geom_histogram()`: 

```{r}
ggplot(mendota, aes(duration)) +
  geom_histogram(fill = "yellow", col = "red") # add bins = 10 to create thicker bars
```


- `geom_density()`: gives the density curve or the (approximated) pdfs.

```{r}
ggplot(mendota, aes(duration)) +
  geom_density()

ggplot(mendota, aes(duration)) +
  geom_density(col = "red", fill = "pink") +
  geom_histogram(mapping = aes(y = after_stat(density)), fill = "yellow", alpha = 0.5)
```

- `geom_boxplot()`: this gives you a distribution and quartile, maxi/min s and outliers


```{r}
ggplot(mendota, aes(duration)) +
  geom_boxplot(fill = "yellow", col  ="red")
```
# Monday 09/21

# Introducing other variables from dataset using the variable aesthetics:

- color or fill

```{r}
ggplot(mendota, aes(duration)) +
  geom_boxplot(mapping = aes(fill = as.factor(decade)))
```

- shapes or sizes for point (linetype and size for line):

```{r}
ggplot(mendota, aes(year1, duration)) + 
  geom_point(mapping = aes(size = as.factor(decade)))
```

```{r}
mpg
```


# Customizations:

### Labeling 

```{r}
ggplot(mendota, aes(year1, duration, col = as.factor(decade))) +
  geom_point() +
  labs(title = "Scatterplot for the duration of lake Mendota closure over the years",
       x = "Year",
       y = "Duration",
       col = "Decade"
       )
```


### Adding reference lines on to plots

- geom_hline(): this adds a horizontal line at the given y-intercept
- geom_vline(): this adds a vertical line at the given x-intercept

```{r}
ggplot(mendota, aes(duration)) +
  geom_histogram() +
  geom_vline(xintercept = mean(mendota$duration), col = "red")
```







