---
title: "ggplot_003"
author: "Sahifa"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE)
```

```{r}
# Assuming that working directory is set to STAT240/Lecture Notes 

library(tidyverse)
mendota = read_csv("../data/lake-mendota-winters-2025.csv")
glimpse(mendota)
```

# GGplot2

GG: Grammar of Graphics

### Step - by - step process

```{r}
# Step 1

ggplot()

# Step 2

ggplot(data = mendota, mapping = aes(x = year1, y = duration))
```

- step 3: `geom_` functions

`geom_point()`: point graph
`geom_line()`: line graph
`geom_histogram()`: create histograms
`geom_density()`: creates a density curve
`geom_bar()`: bar graph


```{r}
# Step 3: layers

ggplot(data = mendota, mapping = aes(x = year1, y = duration)) + 
  geom_point(col = "red") + 
  geom_line(col = "blue")
```


```{r}
ggplot() + 
  geom_point(data = mendota, mapping = aes(x = year1, y = duration), col = "red") + 
  geom_line(col = "blue")
```


# Aesthetic

- classifictation depending on location of the information provided
### Global aesthetic: 
any info provided to `ggplot()` is called a global aesthetic and will affect all `geom_`functions following it.


### Local aesthetic:

any info provided to individual `geom_` functions will only be displayed in that geom_ output.

- depending on type of information

```{r}
ggplot(data = mendota, mapping = aes(x = year1, y = duration)) + 
  geom_point(mapping = aes(col = intervals)) + 
  geom_line(col = "blue")

ggplot(data = mendota, mapping = aes(x = year1, y = duration)) + 
  geom_point(mapping = aes(col = as.factor(intervals))) + 
  geom_line(col = "blue")
```

### Variable aesthetic: 

any information that changes from one point to the other is called a variable aesthetic

### Constant aesthetic:

any information that remains constant throughout the graph is called a constant aesthetic


# Two Variable plots

- `geom_point()` and `geom_line()`

- `geom_smooth()`:

```{r}
ggplot(data = mendota, mapping = aes(x = year1, y = duration)) + 
  geom_point(mapping = aes(col = as.factor(intervals))) +
  geom_smooth(method = "lm", se = F)

ggplot(data = mendota, mapping = aes(x = year1, y = duration)) + 
  geom_point(mapping = aes(col = as.factor(intervals))) +
  geom_smooth(method = "lm", se = F, mapping = aes(col = as.factor(intervals)))
``` 







