---
title: "Joins and pivots (last topic for midterm 1)"
author: "Sahifa"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = T)
library(tidyverse)
```
# Wednesday 09/30

# Joins

A = grocery_list
B = grocery_data


```{r}
A = grocery_list
B = grocery_data

A; B
```

### Mutating Joins

All of these joins add columns from dataset B to dataset A.

`name_join(A, B, by = join_by(column name which is common between both the datasets))`

- 

### Mutating Joins

All of these joins add columns from dataset B to dataset A.

`name_join(A, B, by = join_by(column name which is common between both the datasets))`

Put all of the columns from dataset A frist and then the leftover columns from daatset B.

Dataset A: Rows= 15, Columns= 2 (1 Column is for `item`)
Dataset B: Rows = 12, Columns = 3 (1 Column is for `item`)

Rows: 9 common rows, extra rows in A is 6 (15-9), extra rows in B is 3 (12-9)

Final output of mutating joins: Columns = common column + leftover columns of dataset A + leftover columns of dataset B = 1 + 1 + 2 = 4
 Rows in full join = 9+6+3 = 18


- `left_join()`: Give us all rows that are present in the first dataset

```{r}
left_join(A, B) # rows = 15 (because left dataset which dataset A has 15 rows)
```


- `right_join()`: Give us all rows that are present in the second dataset


```{r}
right_join(A, B) # rows = 12, columns = 4

left_join(B, A)
```

- `inner_join()`: gives the rows that are have same entries in the common column

```{r}
inner_join(A, B) # rows = 9 (same entries from A and B), columns = 4
```


- `full_join()`: gives you all the rows from both of my dataset

```{r}
full_join(A, B) # rows = 18 (9+6+3), columns = 4 
```

### Filtering joins: 

Only give you columns from the first dataset

- `semi_join()`: gives you rows from the first dataset that has a matching entry in the second dataset's common column

```{r}
semi_join(A, B) # rows = 9 (matching rows from A), columns = 2

inner_join(A, B)
```


- `anti_join()`: gives the uncommon rows

```{r}
anti_join(A, B) # rows = 6 (leftover rows from A), columns = 2
```

# Exercise

```{r}
drinks = tibble(
  Row = 1:4,
  Drink = c("Milk Tea", "Espresso", "Latte", "Espresso"),
  Cafe = c("Happy Tea", "Daily Grind", "Daily Grind", "Bev's Beverages"),
  Price = c(4, 2, 3, 2)
)
drinks
```


```{r}
discounts = tibble(
  Row = 1:3,
  Coupon = c("25%", "50%", "5%"),
  Location = c("Happy Tea", "Daily Grind", "The Roast")
)
discounts
```

Dataset A: rows = 4, columns = 4
Dataset B: rows = 3, columns = 3

Common columns: Row, Cafe == Location (2 common cols and 2 extra columns in A and 1 extra column in B)

Common rows: 2 rows (happy tea and daily grind) = 1 extra row in A and 1 extra row in B



- `left_join()`:

```{r}
left_join(drinks, discounts)

left_join(drinks, discounts, by = join_by(Cafe == Location))

left_join(drinks, discounts, join_by(Cafe == Location, Row))
```

# Friday 10/02

# Pivot

### Wide dataset to a long dataset

```{r}
players = tibble(
  player = c("Erling Haaland", "Kevin De Bruyne"),
  goals = c(36, 7),
  assists = c(8, 16),
  games = c(35, 32)
)

players # wide dataset
```


- `pivot_longer()`: pivots wide datasets to long datasets


```{r}
players %>% pivot_longer(c(goals, assists, games))
players %>% pivot_longer(goals:games)
players %>% pivot_longer(- c(player, assists))
players %>% pivot_longer(c(goals,games))


player_l1 =  players %>% pivot_longer(goals:games)
player_l2 = players %>% pivot_longer(-player, names_to = "Characteristic", values_to = "Scores")
```


### Long dataset to a Wide dataset

- `pivot_wider()`:

```{r}
player_l1

player_l1 %>% pivot_wider()

player_l2 %>% pivot_wider(names_from = Characteristic, values_from = Scores)
```

------------
End of Midterm 1 portion

### Removing NA from your dataset

- We have seen that `filter()` combined with `!is.na()` from the dplyr package

- `drop_na()`

```{r}
# starwars %>% filter(!is.na(name), !!is.na(height), !is.na(mass), !is.na(hair_color)......)

starwars %>% drop_na()
starwars %>% drop_na(hair_color)
```


```{r}
table1; table2 

# table1 -> table2, pivot_longer
# table2 -> table1, pivot_wider
```










